MATH 245 Homework 3

### 2024-02-27 PDEs/math245_hw3.Rmd
finished pde hw3

---
title: "MATH 245 Homework 3"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-23"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
##Done: 1, 2, 4, 5, 6.

##Not Done: 3

@@

\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
    \end{cases}       
\end{equation*}

@@

@@

@@

>>>>

---
title: "MATH 245 Homework 3"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-27"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

@@

\begin{equation} \label{eqn:heatIVP}
    \begin{cases}
      v_{t}=k v_{xx} & -\infty <x<\infty,0 <t<\infty \\
      v(0,x)=f(x) & -\infty  <x<\infty \\
    \end{cases}       
\end{equation}

where 
\begin{equation} f(x)=
    \begin{cases}
      x & x>0 \\
      x+1-e^{2x} & x<0 \\
    \end{cases}       
\end{equation}

@@

Let $v(x,t)$ be a solution of $v_x=v_{tt}$. Now, we know that any partial derivatives of a solution to the heat equation are also solutions, and any linear combination of solutions is also a solution. Thus, if $v(x,t)$ is a solution to $v_x=v_{tt}$, so is $w=v_x-2v$. The initial conditions will be

\begin{equation*}
w(x,0)=v_x(0,x)-2v(0,x)= f'(x)-2f(x) = \begin{cases}
   1-2x & \text{ if } x >0 \\
   -1-2x & \text{ if } x<0\\
\end{cases}
\end{equation*}

Thus, $w(x,t)$ satisfies the PDE $v_x=v_{tt}$ with initial conditions $$w(x,0)= \begin{cases}
   1-2x & \text{ if } x >0 \\
   -1-2x & \text{ if } x<0\\
\end{cases}$$

@@

By Problem 1 in this assignment, we know that given the diffusion equation on the whole line with the usual initial condition $u(x, 0)=\phi(x), \phi(x)$ odd, the solution $u(x, t)$ is also an odd function of $x$.

To summarize, we know that
\begin{itemize}
  \item $f'(x)-2f(x)$ is an odd function of $x$
  \item Given a diffusion equation with odd initial conditions, the solution is also odd
  \item $w(x,t)$ is a solution to the diffusion equation with initial conditions $f'(x)-2f(x)$
\end{itemize}

Combining these three points, it is clear that $w(x,t)$ must be an odd function of $x$.
\newline

@@

Since $w(x,t)$ is an odd function, it must automatically satisfy the boundary conditions $w(0,t)=0$. We can solve the problem on the half-line by restricting $x>0$. We know that the solution to 
$$w_t=kw_{xx}, \quad w(x,0)=f'(x)-2f(x), \quad -\infty <x<\infty $$

is given by $$w(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} \left[f'(y)-2f(y) \right]\;dy. $$

Now we can use our solution for $w$ to solve for $v$:
$$v_x-2v= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} \left[f'(y)-2f(y) \right]\;dy $$
$$v_x-2v= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} f'(y)\;dy - 
\frac{2}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} f(y)\;dy$$

We can see that the first term on the right-hand side is the solution to $(vx)_t = k(vx)_{xx}, -\infty<x<\infty$ with the initial condition $v_x(x, 0) = f'(x)$, i.e., it equals $v_x$. Thus, we can subtract $v_x$ from both sides of the above equation and divide by $-2$ to find show that $(\ast)$ is indeed satisfied by
$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy, \quad x>0 $$

Because any solution satisfying the heat equation with specified initial and boundary conditions must be the one and only solution, this must be the unique solution to $(\ast)$.
